---
title: "Introduction"
author: Jonathan Berrisch
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    number_sections: no
    toc: no
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  # dev = "svg",
  warning = TRUE,
  message = FALSE,
  comment = "#>"
)
Sys.setenv("OMP_THREAD_LIMIT" = 2)
```

# A Short Introduction to rcpptimer

This package provides a simple timer for Rcpp code. It is similar to the [tictoc](https://CRAN.R-project.org/package=tictoc) R package. 

The Package wraps [cpptimer](https://github.com/BerriJ/cpptimer) which is a header only library that contains a class called `CppTimer`. 'rcpptimer' adds this class as `Timer` to the `Rcpp` namespace. 

## Basic usage of the Timer with Rcpp::cppFunction

With `Rcpp::cppFunction` we have to add the `depends` argument to the function. To tell the compiler that we want to link the 'rcpptimer' library to the 'Rcpp' code. That is, we can construct an instance if the `Timer` class and use the `tic` and `toc` methods to measure the time it takes to execute a function. Here, we just allocate some memory to have something to measure:

```{r, eval = TRUE}
Rcpp::cppFunction('
int mem()
{
  Rcpp::Timer timer;
  timer.tic("mem");
  std::string s;
  s.reserve(1048576);
  timer.toc("mem");
  return(0);
}',
  depends = "rcpptimer"
)

mem()

print(times)
```

The results will be passed to the R environment as a dataframe named `times`. If you want to give the dataframe a different name you can pass that name to the constructor:

```{r, eval = TRUE}
Rcpp::cppFunction('
int mem()
{
  Rcpp::Timer timer("mytimes");
  timer.tic("mem");
  std::string s;
  s.reserve(1048576);
  timer.toc("mem");
  return(0);
}',
  depends = "rcpptimer"
)

mem()

print(mytimes)
```

## Warnings and how to disable them

The default setting will warn about timers that have not been stopped and `toc` calls for timers that have not yet been started using a matching call to `tic`:

```{r, eval = TRUE}
Rcpp::cppFunction('
int mem()
{
  Rcpp::Timer timer;
  timer.tic("start");
  std::string s;
  timer.tic("reserve");
  s.reserve(1048576);
  timer.toc("reserve");
  timer.toc("finish");
  return(0);
}',
  depends = "rcpptimer"
)

mem()

print(times)
```

Note that this does not affect correctly terminated timers such as `reserve`. These warnings occur at runtime. Unfortunately, we can't check for this at compile time since the `tic` and `toc` calls might spread over various functions so that (in some cases) we do not know the execution flow upfront.

However, if you are sure that you are using the timer correctly you can disable these warnings by passing `false` to the constructor:

```{r, eval = TRUE}
Rcpp::cppFunction('
int mem()
{
  Rcpp::Timer timer(false); // Disable warnings
  timer.tic("start");
  std::string s;
  timer.tic("reserve");
  s.reserve(1048576);
  timer.toc("reserve");
  timer.toc("finish");
  return(0);
}',
  depends = "rcpptimer"
)

mem()
```

## Rcpp::Timer::ScopedTimer

We offer an alternative to the tic-toc interface. The `ScopedTimer` lets you measure the time it takes for the object to go out of scope. We can adjust the above example to use the `ScopedTimer` instead:

```{r, eval = TRUE}
Rcpp::cppFunction('
int mem()
{
  Rcpp::Timer timer;
  Rcpp::Timer::ScopedTimer scoped_timer(timer, "mem");
  std::string s;
  s.reserve(1048576);
  return(0);
}',
  depends = "rcpptimer"
)

mem()

print(times)
```

Note that you only need to initialize the ScopedTimer. Once it goes out of scope it will call `timer.toc("mem")` automatically.

We will add vignettes on how to use the package with Rcpp::sourceCpp and how to add it to your package soon.